Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2014-11-26 manuscript.tex
Filled in missing equations from the methods

A tesseroid is a spherical mass element defined in geocentric coordinates
(Figure~\ref{fig:tesseroid}).
It is bounded by two meridians, two parallels, and two concentric circles.
The gravitational fields of a tesseroid at a point $P = (r,\phi,\lambda)$
are determined with respect to the local North-oriented coordinate system at
P (x, y, z in Figure~\ref{fig:tesseroid}).
Notice that this is a left-handed coordinate system.

@@

\citet{Grombein2013} provides Cartesian kernels for the tesseroid volume
internals that define the gravitational potential, gravitational attraction,
and Marussi tensor, respectively,

\begin{equation}
    V(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{1}{\ell} \kappa  dr' d\phi' d\lambda',
    \label{eq:tesspot}
\end{equation}
\begin{equation}
    g_{\alpha}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{\Delta\alpha}{\ell^3} \kappa dr' d\phi' d\lambda',
    \label{eq:tessgrav}
\end{equation}
\noindent
and
\begin{equation}
    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        I_{\alpha\beta}
        dr' d\phi' d\lambda'.
    \label{eq:tesstensor}
\end{equation}

@@

We refer the reader to \citet{Barrera-Figueroa2006} for an algorithm tailored
for Legendre polynomials in particular.

@@

Thus, the roots must be scaled to the integration limits $[a, b]$ by
\begin{equation}
    x^{scaled}_i = \frac{b - a}{2} x_i + \frac{b + a}{2}.
    \label{eq:glq_scaling}
\end{equation}

The weights of the GLQ are given by \citep{Hildebrand1987},
\begin{equation}
    W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2},
    \label{eq:glq_weights}
\end{equation}

@@

The Gauss-Legendre Quadrature for volume integrals, like equations
\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
becomes \citep{Asgharzadeh2007} (EQUATION).

Applying equation (GLQ 3D) to equations
\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
we see that $f(r_i, \phi_j, \lambda_k)$ is the effect of a point
mass located on the quadrature nodes.

@@

\citet{Li2011} proposed an alternative way to decrease the distance between
point masses and achieve an accurate integration.

>>>>

@@

A tesseroid is a spherical mass element defined in geocentric spherical
coordinates
(Figure~\ref{fig:tesseroid}).
It is bounded by two meridians, two parallels, and two concentric circles.
The gravitational fields of a tesseroid at a point $P = (r,\phi,\lambda)$
are determined with respect to the local North-oriented coordinate system at
P (x, y, z in Figure~\ref{fig:tesseroid}).
\citet{Grombein2013} formulated Cartesian kernels for the volume integrals
that define the tesseroid gravitational potential, gravitational attraction,
and Marussi tensor, respectively,

\begin{equation}
    V(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{1}{\ell}
        \kappa\  dr^\prime d\phi^\prime d\lambda^\prime,
    \label{eq:tesspot}
\end{equation}
\begin{equation}
    g_{\alpha}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \frac{\Delta_\alpha}{\ell^3}
        \kappa\ dr^\prime d\phi^\prime d\lambda^\prime,
    \label{eq:tessgrav}
\end{equation}
\noindent
and
\begin{equation}
    g_{\alpha\beta}(r,\phi,\lambda) = G \rho
        \int\limits_{\lambda_1}^{\lambda_2}
        \int\limits_{\phi_1}^{\phi_2}
        \int\limits_{r_1}^{r_2}
        \left(
            \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
            \frac{\delta_{\alpha\beta}}{\ell^3}
        \right)
        \kappa\ dr^\prime d\phi^\prime d\lambda^\prime.
    \label{eq:tesstensor}
\end{equation}

where $\alpha, \beta \in \{x, y, z\}$,
$\rho$ is the density,
$G = 0.00000000006673\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
$\kappa = {r^\prime}^2 \cos \phi^\prime$,
and

\begin{eqnarray}
    \Delta_x &=& r^\prime\cos\phi\sin\phi^\prime - \sin\phi\cos\phi^\prime
                 \cos(\lambda^\prime - \lambda), \\
    \Delta_y &=& r^\prime \cos \phi^\prime \sin(\lambda^\prime - \lambda), \\
    \Delta_z &=& r^\prime \cos \psi - r,\\
    \ell &=& \sqrt{{r^\prime}^2 + r^2 - 2 r^\prime r \cos \psi}, \\
    \cos\psi &=& \sin\phi\sin\phi^\prime + \cos\phi\cos\phi^\prime
                 \cos(\lambda^\prime - \lambda).
\end{eqnarray}

@@

@@

Before being used for GLQ integration,
the roots must be scaled to the integration limits $[a, b]$ using

\begin{equation}
    x^{scaled}_i = \frac{b - a}{2} x_i + \frac{b + a}{2}.
    \label{eq:glq_scaling}
\end{equation}

The weights of the GLQ are given by \citep{Hildebrand1987},

\begin{equation}
    W_i = \frac{2}{(1 - x_i^2)(P^\prime_N(x_i))^2},
    \label{eq:glq_weights}
\end{equation}

@@

The Gauss-Legendre Quadrature for volume integrals, like equations
\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
becomes \citep{Asgharzadeh2007}

\begin{equation}
    \iiint\limits_{\Omega}
    f(r^\prime, \lambda^\prime, \phi^\prime)
    dr^\prime d\lambda^\prime d\phi^\prime
    \approx
    A
    \sum\limits_{i=1}^{N^r}
    \sum\limits_{j=1}^{N^\phi}
    \sum\limits_{k=1}^{N^\lambda}
    f(r_i, \phi_j, \lambda_k),
    \label{eq:glq3d}
\end{equation}

where

\begin{equation}
    A = \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
\end{equation}

Applying equation~\ref{eq:glq3d} to equations
\ref{eq:tesspot}, \ref{eq:tessgrav}, and \ref{eq:tesstensor},
we see that $f(r_i, \phi_j, \lambda_k)$ is the effect of a point
mass located on the quadrature nodes.

@@

Instead of increasing the GLQ order,
\citet{Li2011} proposed an alternative method
for decreasing the distance between point masses
and achieve an accurate integration.
